miniKanren with fair search strategies

### 2019-05-04 Paper/paper.tex
intro

\newcommand{\conde}{\texttt{cond$^e$} }
\newcommand{\conj}{\texttt{conj}}
\newcommand{\disj}{\texttt{disj}}

@@

miniKanren is a family of relational programming languages. miniKanren 
programs, especially relational interpreters, have been proven to be 
useful in solving many problems by \citet{byrd2017unified}.

A subtlety in writing miniKanren programs with large \conde expressions, such 
as relational interpreters, is that the order of \conde clauses sometimes 
affect the speed considerably. This phenomenon appears when running with the 
miniKanren in \citet{friedman_reasoned_2018}, one of the most well-understood 
implementation. This is because for all \conde expressions with more than two 
clauses, the clauses are given different ``search priority''. Left clauses are 
given higher priorities. This biased treatment causes 2 problems when a \conde 
expression has many clauses: the right-most clause can hardly contribute to 
query result; and the efficiency of the whole program might depend largely on 
the order of these clauses.

@@

In this section, we elaborate fairness by running queries about 
\texttt{repeato}, a relational definition that relates a term \texttt{x} with a 
non-empty list whose elements are \texttt{x} (Fig.~\ref{repeato}).

\begin{figure}
	\lstinputlisting{Figures/repeato.rkt}
	\caption{\texttt{repeato} and an example run}
	\label{repeato}
\end{figure}

>>>>

\newcommand{\conde}{\texttt{cond$^e$} }
\newcommand{\conj}{\texttt{conj}}
\newcommand{\disj}{\texttt{disj}}
\newcommand{\clisting}[1]{
\begin{center}
  \begin{tabular}{c}
	\lstinputlisting{#1}
    \end{tabular}
\end{center}
}

@@

miniKanren is a family of relational programming languages.
\citet{friedman_reasoned_2018} introduce miniKanren and its implementation
in the The Reasoned Schemer, 2nd Ed(TR2). miniKanren programs, especially
relational interpreters, have been proven to be 
useful in solving many problems by \citet{byrd2017unified}.

\conde is the flagship disjunction operator in miniKanren. A subtlety arises
when a \conde contains many disjunction clauses: not every clause has the equal
chance to contribute to the answers. As an example, consider the following
relation \texttt{repeato} and its invocation.
\clisting{Figures/repeato.rkt}
Next, consider the following disjuction of invoking \texttt{repeato} with
four different letters.
\clisting{Figures/example.rkt}
\conde intuitively corresponds to \texttt{or} in logic. An unsuspicious beginner
would expect each letter to contribute equally to the result, as following. 
\clisting{Figures/run-repeato-fair.rkt}
The \conde in TR2, however, generates a less expected result.
\clisting{Figures/run-repeato-idfs.rkt}

A subtlety arises when 
A subtlety in writing miniKanren programs with \conde expressions with many
clauses, such 
as relational interpreters, is that the order of \conde clauses sometimes 
affect the speed considerably. This phenomenon appears when running with the 
miniKanren in \citet{friedman_reasoned_2018}, one of the most well-understood 
implementation. This is because for all \conde expressions with more than two 
clauses, the clauses are given different ``search priority''. Left clauses are 
given higher priorities. This biased treatment causes 2 problems when a \conde 
expression has many clauses: the right-most clause can hardly contribute to 
query result; and the efficiency of the whole program might depend largely on 
the order of these clauses.

@@

In this section, we elaborate fairness by running queries about 
\texttt{repeato2}, a relational definition that relates a term \texttt{x} with a 
non-empty list whose elements are \texttt{x} (Fig.~\ref{repeato}).

\begin{figure}
	\lstinputlisting{Figures/repeato.rkt}
	\caption{\texttt{repeato} and an example run}
	\label{repeato2}
\end{figure}
